---
runtime: shiny
output:
  html_document:
    theme: paper
    highlight: tango
    code_fold: show
---

```{css, echo=FALSE}
footer {
  min-height: 100px;
  background-color: #323e48;
  color: #e0e2e4;
}

.nav-wrapper {
  background-image: linear-gradient(135deg, #1a4787 0%, #0891bf 100%);
  height: 90px;
  padding: 12px 0;
  width: 100%; 
  z-index: 2;
}

.brand-logo {
  position: absolute;
  color: #fff;
  font-size: 2.1rem;
  padding: 0;
}

.rds-logo { 
  margin-left: 8px;
  margin-top: 2px;
  max-width: 190px;
}

.mtna-logo {
  margin-left: 8px;
  margin-top: 4px;
  max-width: 190px;
}

.right{
  float:right;
}

.nav-tabs>li>a{
  font-size: 20px;  
}

```

<div class="nav-wrapper">
  <a href="https://www.richdataservices.com/" class="brand-logo rds-logo"><img src="./RDS-Logo_DARK.svg" class="rds-logo"></a>
  <h1 class="right" style="color:white;">COVID-19 Examples</h1>
</div>

#### Getting Started

<div>
  Don't know where to begin? The following resources can help get you started working with our COVID-19 data.
  <ul>
    <li><a href="https://covid19.richdataservices.com/" target="_blank" >RDS COVID-19 website</a> - introduction to the tools and services Metadata Technology North America (MTNA) provides for accessing COVID-19 data.</li>
    <li><a href="https://github.com/mtna/covid-19" target="_blank">RDS COVID-19 GitHub repository</a> - the source data files use in the RDS COVID-19 catalog along with information about the sources we pull from and the transformations that were performed on them.</li>
    <li><a href="https://github.com/mtna/rds-r" target="_blank">RDS R GitHub repository</a> - the R library for interacting with RDS.</li>
  </ul>
</div>

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(rds.r)
library(ggplot2)
```

#### Using COVID-19 Data with RDS

There are a variety of data products available through the RDS API around COVID-19 and it is important to know what RDS methods to use based on the data that is included in the data product. This will go through two examples explaining the differences between some of the data products available and what to look for when querying the data. 

First, connect to the RDS server and get the COVID-19 catalog that houses the data products. 

```{r }
# factors vs data types
# connect to the rds server
rds <- get.rds("http://dev.richdataservices.com")

# get the covid catalog
catalog <- getCatalog(rds, "covid19")
```

## {.tabset}

### Time Series Data

#### Johns Hopkins University Country Data

Many of the COVID data products being collected are time series that contain aggregate counts for certain dimensions. The dimensions will always include a date and then any others that are being measured such as country, state, county, etc. For time series the `select` method will be a good place to start. 

The country level data collected by the Johns Hopkins University Center for Systems Science and Engineering is a good example of time series data. It contains the counts of confirmed cases, active cases, deaths, and recoveries by date and country. 

RDS will empower us to quickly create a country selector and run the queries to look at the counts based on the selected country.

##### Getting the Data Product

The JHU CCSE Country data product can be retrieved from the COVID-19 catalog that has already been retrieved.

```{r }
# get the JHU Product
jhuCountry <- getDataProduct(catalog, "us_jhu_ccse_country")
```

##### Discovering Countries to View

So what countries are available? The country variable `iso3166_1` is a classified variable in this data set. This means that RDS provides the entire code list of valid countries which can be used set up the select input.

```{r }
# get the classification associated with the date variable, iso3166_1
countries <- getClassification(jhuCountry, "iso3166_1")

# use the first country returned as the initial value in the select input
firstCountry <- countries[,"name"][1]

# the select input will be rendered close to the graph
```

##### Querying the Data

With the country codes retrieved it is time to set up a method to query the data product and get the data to plot. In this example the confirmed case, death, and recovery counts will be queried for each country. 

Because this is a time series data product it is critical that the `date_stamp` variable is included to get the counts over time.

```{r }
# set up method to retrieve the data using the input country
select <- reactive({
  
  # get the associated country code (input$country) from the country name the user selected
  countryCode <- countries[countries$name == input$country, "codeValue"]
 
  # get the data set with the date, counts of confirmed, death, recovered filtered by country 
  dataSet <- rds.select(
    jhuCountry,
    cols = c("date_stamp", "cnt_confirmed", "cnt_death", "cnt_recovered"),
    where = paste("iso3166_1=", countryCode, sep = "", collapse = NULL)
  )
  
  dataSet@records
})
```

##### Plotting the Data

Now the country select input can be set up, the input will be captured and used in the querying method to retrieve the data frame to use and the chart can be plotted.

```{r }
  # set up the select input for users to pick a country (using the countries$name column)
  selectInput("country", label = "Country:",
                choices = countries$name, selected = firstCountry)

  renderPlot({
    df <-select()
    shiny::validate(need(nrow(df) > 0, paste("No data found for", input$country)))
    ggplot(df, aes(x = date_stamp, y = cnt_confirmed)) +
      scale_x_date(date_breaks  = "1 week") +
      geom_line(aes(colour = "Confirmed Cases"), group = 1) +
      geom_line(aes(y = cnt_death, colour = "Deaths"), group = 1) +
      geom_line(aes(y = cnt_recovered, colour = "Recovered"), group = 1) +
      xlab("") + ylab("Count") +
      scale_colour_manual(name = "Count of:",
                          values = c("blue", "red", "green")) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  })
```


### Record Level Data

#### Statistics Canada Data

A number of COVID data products, as well as the contextual survey data that are available through RDS, are record level data sets. This means that each record in the data product measures characteristics or and information of a single individual or respondent. In these cases to create charts and graphs the RDS `tabulate` method can be used to group measures by the desired dimensions. 

The confirmed cases collected and distributed by Statistics Canada is a good example of record level data, as it measures the latest hospitalization, ICU, and deceased status of individuals confirmed to have coronavirus on a specific date.

This example will build a graph about the individuals by gender and age group and allow the customization of the graph through the selection of transmission type, hospitalization, and death status.

##### Getting the Data Product

The Statistics Canada data product can be retrieved from the COVID-19 catalog that has already been retrieved.

```{r }
# get the Statistics Canada Product
statCan <- getDataProduct(catalog, "ca_statcan")
```

##### Building the Filtering Selections

The variables we will be using to filter the data, `transmission_type`, `is_hospitalized`, and `is_deceased`, are all categorical variables and their code lists are available through the RDS API. We will use these to populate our select inputs. We will add an additional value of "All" which will be the starting value. 

```{r }
# add a default value which means do not filter on this type or status
default <- "All"

# get the classifications associated with the variables used to filter
# set up the options for users to pick from the default value or code names associate with the variable

# transmission type
transmissionType <- getClassification(statCan, "transmission_type")
transmissionChoices <- c(default)
transmissionChoices <- append(transmissionChoices, transmissionType$name)

# is hospitalized
isHospitalized <- getClassification(statCan, "is_hospitalized")
hospitalizedChoices <- c(default)
hospitalizedChoices <- append(hospitalizedChoices, isHospitalized$name)

# is deceased
isDeceased <- getClassification(statCan, "is_deceased")
deceasedChoices <- c(default)
deceasedChoices <- append(deceasedChoices, isDeceased$name)

# the select input for all of these will be rendered close to the graph
```

##### Querying the Data

With the codes retrieved it is time to set up a method to query the data product and get the data to plot. In this example the tabulate method will be used to get the count of cases by age group and gender. 

```{r }
# set up method to retrieve the data using the inputs by the select boxes
tabulation <- reactive({
  
  # build up the where clause
  where <- c()
  if (input$transmission != default) {
    where <-
      append(where, paste(
        "transmission_type=",
        transmissionType[transmissionType$name == input$transmission, "codeValue"],
        sep = ""
      ))
  }
  if(input$hospitalized != default) {
    where <-
      append(where, paste(
        "is_hospitalized=",
        isHospitalized[isHospitalized$name == input$hospitalized, "codeValue"],
        sep = ""
      ))
  }
  if (input$deceased != default) {
    where <-
      append(where, paste(
        "is_deceased=",
        isDeceased[isDeceased$name == input$deceased, "codeValue"],
        sep = ""
      ))
  }
  if (length(where) > 0){
    where <- paste(where, collapse = " AND ")
  }
  
  # run the tabulation to get the count of individuals by gender and age group
  dataSet <-
    rds.tabulate(
      statCan,
      dimensions = c("gender", "age_group"),
      inject = T,
      where = where
    )
  dataSet@records
})
```

##### Plotting the Data

Now the select inputs can be set up, the input will be captured and used in the querying method to retrieve the data frame to use and the graph can be plotted.

```{r }
# Create the inputs and display them side by side
div(
  div(
    style = "display:inline-block",
    selectInput(
      "transmission",
      label = "Transmission Type:",
      choices = transmissionChoices,
      selected = default
    )
  ),
  div(
    style = "display:inline-block",
    selectInput(
      "hospitalized",
      label = "Is Hospitalized:",
      choices = hospitalizedChoices,
      selected = default
    )
  ),
  div(
    style = "display:inline-block",
    selectInput(
      "deceased",
      label = "Is Deceased:",
      choices = deceasedChoices,
      selected = default
    )
  )
)

# plot the graph
renderPlot({
  ggplot(tabulation(), aes(x = gender, y = age_group, fill = age_group, size = count)) + 
    scale_size_continuous(range=c(1,20)) +
    geom_point(shape = 21) + xlab("Gender") + ylab("Age Group") + labs(fill = "Age Group", size = "Count")
})
```

##